---
title: "ILINet Forecast"
output: html_notebook
---
## Setting up libraries
```{r}
library(forecast)
library(lubridate)
library(tidyverse)
library(prophet)
library(Metrics)

```
## Loading Train/Test data
```{r}
train <- readRDS('data/train.rds')
test  <- readRDS('data/test.rds')
```
## Data split by state for individual predictions
Assuming that different States have different behaviour regarding Flu cycles because of latitude and longitude differences, we'll fit a model to each one.
```{r tidy=FALSE}
data_split <- function(data){
    
  data_sp <- lapply(
    unique(data$REGION), 
    function(x) data[data$REGION == x,]
  )
}

test_split  <- data_split(test)
train_split <- data_split(train)
```


## Forecasting
The Prophet library takes as input a dataframe with two variables: **ds** for the dates in the format YYYY-MM-DD and **y** for the numeric value.

```{r echo=T, results='hide', tidy=FALSE}
models_profet    <- list()
forecasts_profet <- list()
for(state_df in train_split){
  model <-state_df %>% 
    select(ds = PRETTYDATE, y = ILITOTAL) %>% 
    prophet(
      yearly.seasonality = TRUE, 
      weekly.seasonality = TRUE, 
      daily.seasonality = FALSE
    )
  
  future <- make_future_dataframe(
    model, 
    periods = 4, 
    freq = 'week'
  )
  pred <- predict(model, future)

  models_profet[unique(state_df$REGION)] <- list(model)
  forecasts_profet[unique(state_df$REGION)] <- list(pred)
}

```


## Results (This function plots for the selected state)
```{r}
plot_forecast_prophet <- function(state){
  plot(
    models_profet[[state]], 
    forecasts_profet[[state]]
  )
}
```

```{r}
plot_forecast_prophet('New York')
```


## Forecasting using auto-arima since configure ARIMA one by one is difficult 
ARIMA 

```{r echo=T, results='hide', tidy=FALSE}
models_auto_arima    <- list()
forecasts_auto_arima <- list()
counter <-1
for(state_df in train_split){
  model <-
    auto.arima(
        ts(state_df$ILITOTAL,
           frequency=52,
           start=decimal_date(ymd("2010-10-04")))
        )
  
  pred <- forecast(model,h=4)
    
  models_auto_arima[unique(state_df$REGION)] <- list(model)
  forecasts_auto_arima[unique(state_df$REGION)] <- list(pred)
}
```

```{r}
saveRDS(models_auto_arima, 'data/models_arima.rds')
saveRDS(forecasts_auto_arima, 'data/forecasts_arima.rds')
```

```{r}
models_auto_arima <- readRDS('data/models_arima.rds')
forecasts_auto_arima <- readRDS('data/forecasts_arima.rds')
```

## Results (This function plots for the selected state) for Arima , break is something ugly to do. 

##Try to plot it using funggcast but it doesn't support weekly ts

```{r}
plot_forecast_auto_arima <- function(state){
  for(state_df in train_split){
    if(state_df$REGION==state){
      temporal_state_df = state_df
      break
    }
  }
   plot(ts(temporal_state_df$ILITOTAL,frequency=52, start=c(2010,40), end=c(2018,20)))
   lines(fitted(models_auto_arima[[state]]), col='red')
      
}
plot_forecast_auto_arima('Colorado')


```

##Accuracy using test data -  4 weeks is insufficient to be significant 


```{r}

mase_dict_test <- list()
counter = 1;
for (state_name in names(models_auto_arima)){
  predicted_profet <-mase(test_split[[counter]]$ILITOTAL,tail(forecasts_profet[[state_name]]$yhat,4))
  predicted_auto_arima <-mase(test_split[[counter]]$ILITOTAL,tail(forecasts_auto_arima[[state_name]]$fitted,4))
  
  if(predicted_profet < predicted_auto_arima){
    mase_dict_test[[paste(state_name, "profet", sep="-")]]<- c( predicted_auto_arima ,predicted_profet)
  }else{
    mase_dict_test[[paste(state_name, "arima", sep="-")]] <-c( predicted_auto_arima ,predicted_profet)
  }
  counter = counter +1
}
mase_dict_test

```

##Accuracy using train data 
 
```{r}

mase_dict_train <- list()
counter = 1;
for (state_name in names(models_auto_arima)){
  lenghtOfSet = length(forecasts_profet[[state_name]]$yhat)
  predicted_profet <-mase(train_split[[counter]]$ILITOTAL,head(forecasts_profet[[state_name]]$yhat,lenghtOfSet-4))
  predicted_auto_arima <-mase(train_split[[counter]]$ILITOTAL,head(forecasts_auto_arima[[state_name]]$fitted,lenghtOfSet-4))
  
  if(predicted_profet < predicted_auto_arima){
    mase_dict_train[[paste(state_name, "profet", sep="-")]]<- c( predicted_auto_arima ,predicted_profet)
  }else{
    mase_dict_train[[paste(state_name, "arima", sep="-")]] <- c( predicted_auto_arima ,predicted_profet)
  }
  counter = counter +1
}
mase_dict_train

```



```{r include=FALSE}
rm(future, model, pred, state_df, test, train, actual, predicted_auto_arima, predicted_profet, x,y); gc()
```