---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(magrittr)
library(plotly)
library(forecast)
library(ggplot2)
```

# Data Loading ----
# Working with ilinet only as the table broke by age group doesn't have State info

```{r}
ilinet_tbl  <- read_csv('input/ILINet.csv') %>% 
  arrange(REGION, YEAR, WEEK)
```

Preprocessing ----
# Check number of unique values per feature

```{r}
map(ilinet_tbl, ~length(unique(.x)))
```

# Drop irrelevant features
```{r}
ilinet_tbl <- ilinet_tbl[, map_lgl(ilinet_tbl, ~length(unique(.x)) > 2)]
```

# Investigate why WEEK has 53 distinct values (a year has 52 weeks)
```{r}
table(ilinet_tbl$WEEK)
```

# TODO verify what is the week 53 (maybe leap year?), dropping for now
```{r}
ilinet_tbl %<>%
  filter(WEEK != 53)
```

# Coerce numerical features to numeric (this will introduce NAs where value == 'X')
```{r}
ilinet_tbl[, 4:7] <- map(ilinet_tbl[, 4:7], as.numeric)
```

# Check number of NAs per feature
```{r}
ilinet_tbl %>% 
  group_by(REGION) %>% 
  summarise_all(function(x){sum(is.na(x))}) %>% 
  View
```
# NAs are mostly divided between Florida and Virgin Islands, let's
# investigate further
```{r}
ilinet_tbl %>% 
  filter(REGION %in% c('Florida', 'Virgin Islands')) %>% 
  group_by(REGION) %>% 
  summarise_all(function(x){sum(is.na(x))/length(x)}) %>% 
  View
```

# 100% of Florida info is NA, 60% of Virgin Islands info is NA
# TODO decide what to do with these states later, dropping for now
```{r}
ilinet_tbl %<>% 
  filter(!(REGION %in% c('Florida', 'Virgin Islands')))
```
# Combine YEAR and WEEK info

```{r}
ilinet_tbl %<>%
  mutate(
    YEARWEEK   = paste(YEAR, WEEK, sep = '-'), 
    PRETTYDATE = as.Date(paste(YEAR, WEEK, 1, sep="-"), "%Y-%U-%u")
  )
```
# Train/Test split ----
# Use last 3 weeks as Test Set (as data is released w/ 2 weeks lag + 1 week protection
# against data release delay)
```{r}
train <- ilinet_tbl %>% 
  filter(PRETTYDATE < max(PRETTYDATE) - lubridate::weeks(3))

test <- ilinet_tbl %>% 
  filter(PRETTYDATE >= max(PRETTYDATE) - lubridate::weeks(3))
```

```{r}
 saveRDS(train, 'data/train.rds')
 saveRDS(test, 'data/test.rds')
 write_csv(train, 'data/train.csv')
 write_csv(test, 'data/test.csv')
```

#General overview of some of the regions
```{r}
plot_ly(train, x = ~REGION, y = ~ILITOTAL,  type = "box")
```


```{r}
trainSubSet <- filter(train, REGION %in% c('Alabama', 'California','Hawaii','New York','Virginia'))
sp <- ggplot(trainSubSet, aes(x=trainSubSet$PRETTYDATE, y=trainSubSet$ILITOTAL)) + geom_line()
sp + facet_grid(unique(trainSubSet$REGION) ~ .)
```


```{r}
trainSubSet <- filter(train, REGION %in% c('New Mexico', 'Louisiana','Montana','Oklahoma','Vermont'))
sp <- ggplot(trainSubSet, aes(x=trainSubSet$PRETTYDATE, y=trainSubSet$ILITOTAL)) + geom_line()
sp + facet_grid(unique(trainSubSet$REGION) ~ .)
```

